Lightweight Transfer Learning from Knowledge Base to Microblog Stream for Accurate Event Detection

### 2016-09-15 paper/aaaiFormat.tex
push sec 4.1 forward

%在Related Works 之前说明我们这是一个amazing的工作，同时将precision和recall都进行了提升，而一般的工作很难兼顾这一点。
But existing methods cannot achieve high precision and high recall simultaneously. 
The reason lies in that they only use the recent data and lack the accumulation of history knowledge. 
We further find out that the ideal event detection method should have two properties: \textit{organizing history data well} and \textit{self-updating}.
According to this guideline, we propose \textsc{NSDetector}, a high accuracy event detection method that trains normal states better by using knowledge base, which can be treated as the accumulation of history data. 
In \textsc{NSDetector}, 1) a new data structure \textit{normal states} is proposed to restore what happened; 2) \textit{normal states} can be initialized by knowledge base and further updated by upcoming text streams via our proposed probabilistic model NS-Prior-LDA; 3) events can be detected by normal states with high accuracy.
Experiments on the benchmark \textit{Edinburgh twitter corpus (30 million tweets)} validate the effectiveness of our proposed \textsc{NSDetector}.
%\textsc{NSDetector} imitates this process automatically in the event detection task, and further verify that the key to improve the accuracy of event detection lies in training the normal states well. 
\end{comment}
\end{abstract}

%\textsc{NSDetector} imitates this process automatically in the event detection task, and further verify that the key to improve the accuracy of event detection lies in training the normal states well.

\section{Introduction}
Many web applications need the accurate event detection technique on microblog stream, such as public opinion analysis\cite{thelwall2011sentiment}, public security\cite{Li:2012gw}, and disaster response\cite{Yin:2012ht}, etc.
Although event detection has been a research topic for a long while\cite{allan1998topic}, but event detection in microblog stream is still challenging\cite{atefeh2015survey}.
According to \cite{huang2016probabilistic}, the characteristics of microblog is fast changing, high noise and short length.

Different from the not-well-organized microblog stream, knowledge base (e.g. Wikipedia) is constructed elaborately and contains rich information. 
For this reason, knowledge base can be a good supplementary for event detection on microblog stream.
For example, the tweet ``''

But it's non-trivial to perform transfer learning from knowledge base to microblog stream.

@@

The properties of the ideal event detection method also coincide with the news value theory\cite{galtung1965structure}\cite{caple2013delving}, which emphasizes the \textit{news unexpectedness} -- ``the unpredictable or the rare is more newsworthy than the routine"\cite{bell1991language}, and ``routine events are difficult to assimilate to the news, which favours the novel, the atypical and the unusual"\cite{montgomery2007discourse}.
Following this guideline, if the journalist knew the normal states, aka, the ``routine things" of everyday better, he/she would tend to choose more unexpected news to report. 
We suppose that the ideal event detection system performs like a journalist who can find news from cyberspace with the knowledge of the normal states. 
%Taking another example in our study for example, the tweet ``UN declares famine in Somalia (7/20/2011)"\footnote{\url{https://en.wikipedia.org/wiki/2011_East_Africa_drought}} is related to a \textit{public security} event as the famine has the potential of relating to a war.
%Being aware of the history knowledge, such as \textit{Vietnamese Famine of 1945}\footnote{\url{https://en.wikipedia.org/wiki/Vietnamese_Famine_of_1945}}, \textit{Dutch famine of 1944-45}\footnote{\url{https://en.wikipedia.org/wiki/Dutch_famine_of_1944-45}} caused by World War II, it's possible to detect it accurately.

@@

\begin{figure}[h]
    \includegraphics[width=1.0\columnwidth]{img/NSDetectorExample.pdf}
    \caption{\textsc{NSDetector}'s process flow, taking \textit{public security} related events in cyberspace as an example. \textsc{NSDetector} initializes the normal states about military from knowledge base, learns the topical words further within the time windows on the target corpus, detects events, and updates the normal states incrementally by the detected events.}
    \label{fig:modelDesc}
\end{figure}

@@

\textbf{History States and Recent States.} 
The history states are initialized on the pre-defined categories on the knowledge base as Algorithm \ref{alg:normalStatesInit}.
For \textsc{HS-Prior-LDA}, \(K\) is set to be 200, which means \textsc{HS-Prior-LDA} learned 100 history-state-like topics and 100 other topics.
After cross-validated grid-search, \(\lambda\) is set to be 12.8. The other parameters are set as \(\alpha=0.1\), \(\beta=0.005\).
We run \textsc{HS-Prior-LDA} window by window on text stream, and learn specific categories' recent states.
This process is illustrated by Table \ref{tbl:historyStates}.

We compare the topic coherence of history states with the topics learned from Wikipedia by LDA in terms of NPMI\cite{Rder2015ExploringTS}.
Different from traditional experiments that only compute the topic coherence of top words, we want to check whether  it can hold for more words.
Due to the limit of NPMI computing module\footnote{\url{https://github.com/AKSW/Palmetto}}, which computes the coherence of up to 10 words each time, we divide the top 100 words into 20 groups.

@@

%\begin{comment}%TODO
\begin{figure}
        \centering
        \includegraphics[width=1.0\columnwidth]{img/NPMI.pdf}
        \caption{The comparison on NPMI between LDA's topics and history states initialized from Wikipedia.}
\end{figure}
%\end{comment}

@@

\begin{table*}[ht]
\centering
\caption{History States initialized from Wikipedia and topical words learned from HS-Prior-LDA}
\scalebox{0.76}{
\begin{tabular}{|cc|cc|cc|cc|cc|cc|}
\hline
\multicolumn{2}{|c|}{\textit{Aviation}} & \multicolumn{2}{c|}{\textit{Health}} & \multicolumn{2}{c|}{\textit{Middle East}} & \multicolumn{2}{c|}{\textit{Military}} & \multicolumn{2}{c|}{\textit{Mobile Phones}} & \multicolumn{2}{c|}{\textit{Video Games}} \\
\begin{tabular}[c]{@{}c@{}}Normal\\ States\end{tabular} & \begin{tabular}[c]{@{}c@{}}Topical\\ Words\end{tabular} & \begin{tabular}[c]{@{}c@{}}Normal\\ States\end{tabular} & \begin{tabular}[c]{@{}c@{}}Topical\\ Words\end{tabular} & \begin{tabular}[c]{@{}c@{}}Normal\\ States\end{tabular} & \begin{tabular}[c]{@{}c@{}}Topical\\ Words\end{tabular} & \begin{tabular}[c]{@{}c@{}}Normal\\ States\end{tabular} & \begin{tabular}[c]{@{}c@{}}Topical\\ Words\end{tabular} & \begin{tabular}[c]{@{}c@{}}Normal\\ States\end{tabular} & \begin{tabular}[c]{@{}c@{}}Topical\\ Words\end{tabular} & \begin{tabular}[c]{@{}c@{}}Normal\\ States\end{tabular} & \begin{tabular}[c]{@{}c@{}}Topical\\ Words\end{tabular} \\ 
\hline
aircraft & air\scriptsize(3253) & health & weight\scriptsize(16344) & al & \textcolor{red}{\textit{\#syria}}\scriptsize(4212) & army & killed\scriptsize(4055)  & android & iphone\scriptsize (13674) & game & games\scriptsize(8812) \\ 
air & plane & patients & loss & israel & \textcolor{red}{\textit{\#bahrain}} & military & news & mobile & apple & player & liked \\ 
airport & flight & medical & diet & iran & people & air & \textcolor{red}{\textit{\#libya}}\scriptsize(3503) & nokia & android & playstation & free \\ 
flight & time & disease & health & arab & israel & command & libya & ios & app & gameplay & xbox \\
airline & airlines & treatment & cancer & israeli & police & force & rebels & phone & ipad & nintendo & 360 \\
airlines & news & hospital & lose & egypt & \textcolor{red}{\textit{\#libya}}\scriptsize(2557) & regiment & people & samsung & samsung & games & playing \\
aviation & boat & patient & fat & egyptian & \#egypt & forces & police & game & mobile & players & played \\
flying & airport & clinical & tips & ibn & news & squadron & war & app & blackberry & xbox &iphone\scriptsize(2820)\\
pilot & force & symptoms & treatment & jerusalem & \#israel & infantry & libyan & iphone & tablet & mode & time\\
squadron & fly & cancer & body & syria & world & battle & attack & htc & apps & arcade & \textcolor{red}{\textit{ps3}}\\
\hline
\end{tabular}
}
\label{tbl:historyStates}
\end{table*}

\begin{comment}
\begin{table*}[ht]
\centering
\caption{Event Detection Results on Edinburgh Twitter Corpus}
\label{tbl:edinburgh_results}
\subtable[Edinburgh Twitter Corpus Statistics]{
\begin{tabular}{|l|r|}
\hline
Time range & \begin{tabular}[c]{@{}l@{}}2011/06/30\\ -2011/09/15\end{tabular} \\ \hline
Tweet Number & 36,627,434 \\ \hline
\begin{tabular}[c]{@{}l@{}}Event Number \\ in Benchmark1\end{tabular} & 27 \\ \hline
\begin{tabular}[c]{@{}l@{}}Event Number \\ in Benchmark2\end{tabular} & 523 \\ \hline
\end{tabular}
}\quad \quad
\subtable[Event Detection Results]{
    \begin{tabular}{|l|l|l|l|l|}
    \hline
    Method & \begin{tabular}[c]{@{}l@{}}Recall@ \\ Benchmark1\end{tabular} & \begin{tabular}[c]{@{}l@{}}Precision@ \\ Benchmark2\end{tabular} & \begin{tabular}[c]{@{}l@{}}Recall@ \\ Benchmark2\end{tabular} & \begin{tabular}[c]{@{}l@{}}DERate(Duplicate \\ Event Rate)\end{tabular} \\ \hline
    UMass System\cite{Allan:2000wu} & 0.882 & 0.138 & 0.941 & 0.071 \\ \hline
    LSH & 0.824 & 0.095 & 0.803 & 0.302 \\ \hline
    TimeUserLDA & 0.353 & 0.536 & 0.071 & 0.054 \\ \hline
    Twevent & 0.824 & 0.697 & 0.641 & 0.113 \\ \hline
    EDCoW & 0.412 & 0.756 & 0.119 & 0.290 \\ \hline
    BurstyBTM & 0.647 & 0.809 & 0.170 & 0.045 \\ \hline
    \textsc{NSDetector} & 1.000 & 0.894 & 0.950 & 0.042 \\ \hline
    \end{tabular}
}
\end{table*}
\end{comment}

\begin{comment}
\begin{table*}[ht]
\centering
\caption{Events about \textit{vehicles} detected by systems between 2011-07-20 and 2011-07-30}
\label{my-label}
\begin{tabular}{|l|L{3cm}|l|c|}
\hline
\multicolumn{1}{|c|}{Date} & \multicolumn{1}{c|}{Event key words} & \multicolumn{1}{c|}{Representative event tweet} & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Number of \\ event tweet\end{tabular}} \\ \hline
7/20/11 & \begin{tabular}[c]{@{}l@{}}American, airlines, \\ order, planes\end{tabular} & \begin{tabular}[c]{@{}l@{}}NDTV: American Airlines orders 460 new planes from \\ Boeing, Airbus http://bit.ly/p1ZgYG\end{tabular} & 31 \\ \hline
7/26/11 & \begin{tabular}[c]{@{}l@{}}Morocco, military, \\ plane, crash\end{tabular} & \begin{tabular}[c]{@{}l@{}}BBC News - Morocco military plane crash kills 78 \\ http://t.co/uwnLv8L\end{tabular} & 36 \\ \hline
7/27/11 & air, Canada & \begin{tabular}[c]{@{}l@{}}Smoke in galley forces Air Canada flight back to airport. \\ Plane landed safely in Sydney, Australia after dumping fuel\end{tabular} & 48 \\ \hline
7/30/11 & \begin{tabular}[c]{@{}l@{}}Caribbean, airlines, \\ crashes, Guyana\end{tabular} & \begin{tabular}[c]{@{}l@{}}Caribbean Airlines Jet Crash-Lands In Guyana All 163 \\ Passengers, Crew Survive http://t.co/qPVD3WJ\end{tabular} & 57 \\ \hline
\end{tabular}
\end{table*}
\end{comment}

>>>>

%在Related Works 之前说明我们这是一个amazing的工作，同时将precision和recall都进行了提升，而一般的工作很难兼顾这一点。
But existing methods cannot achieve high precision and high recall simultaneously. 
The reason lies in that they only use the recent data and lack the accumulation of history knowledge. 
We further find out that the ideal event detection method should have two properties: \textit{organizing history data well} and \textit{self-updating}.
According to this guideline, we propose \textsc{NSDetector}, a high accuracy event detection method that trains normal states better by using knowledge base, which can be treated as the accumulation of history data. 
In \textsc{NSDetector}, 1) a new data structure \textit{normal states} is proposed to restore what happened; 2) \textit{normal states} can be initialized by knowledge base and further updated by upcoming text streams via our proposed probabilistic model NS-Prior-LDA; 3) events can be detected by normal states with high accuracy.
Experiments on the benchmark \textit{Edinburgh twitter corpus (30 million tweets)} validate the effectiveness of our proposed \textsc{NSDetector}.
\end{comment}
\end{abstract}

\section{Introduction}
Many web applications need the accurate event detection technique on microblog stream, such as public opinion analysis\cite{thelwall2011sentiment}, public security\cite{Li:2012gw}, and disaster response\cite{Yin:2012ht}, etc.
Although event detection has been a research topic for a long while\cite{allan1998topic}, event detection in microblog stream is still challenging\cite{atefeh2015survey}.
According to \cite{huang2016probabilistic}, the characteristics of microblog, which is fast changing, high noise, and short length, raise the challenge.

Knowledge base can be a good supplementary for event detection on microblog stream.
Different from the not-well-organized microblog stream, knowledge base (e.g. Wikipedia) is constructed elaborately and contains rich information. 
For example, the microblog message ``libyan rebel chief gunned down in Benghazi (2011-07-28)'' is short, but still comprehensible because the words ``\textit{libyan}", ``\textit{Benghazi}'' are included in the wiki page ``\textit{Libya}", and the words ``\textit{rebel}'', ``\textit{gun}'' appeared in the wiki page ``\textit{Military}''.
By reading these two wiki pages, readers easily understand the example tweet is talking about something related to \textit{Military} and \textit{Libya}.
In other words, knowledge base enrichs the linkage
Since transfer learning aims at , it provides a way to .

But it's non-trivial to perform transfer learning directly from knowledge base to microblog stream.

@@

The properties of the ideal event detection method also coincide with the news value theory\cite{galtung1965structure}\cite{caple2013delving}, which emphasizes the \textit{news unexpectedness} -- ``the unpredictable or the rare is more newsworthy than the routine"\cite{bell1991language}.
Following this guideline, if the journalist knew the normal states, aka, the ``routine things" of everyday better, he/she would tend to choose more unexpected news to report. 
We suppose that the ideal event detection system performs like a journalist who can find news from cyberspace with the knowledge of the normal states.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=.82\columnwidth]{img/NSDetectorExample.pdf}
    \caption{\textsc{LTDetector}'s process flow, taking \textit{Military} related events in cyberspace as an example. \textsc{LTDetector} initializes the normal states about military from knowledge base, learns the topical words further within the time windows on the target corpus, detects events, and updates the normal states incrementally by the detected events.}
    \label{fig:modelDesc}
\end{figure}

@@

\textbf{History States and Recent States.} 
The history states are initialized on the pre-defined categories on the knowledge base as Algorithm \ref{alg:normalStatesInit}.
For \textsc{HS-Prior-LDA}, \(K\) is set to be 200, which means \textsc{HS-Prior-LDA} learned 100 history-state-like topics and 100 other topics.
After cross-validated grid-search, \(\lambda\) is set to be 12.8. The other parameters are set as \(\alpha=0.1\), \(\beta=0.005\).
We run \textsc{HS-Prior-LDA} window by window on text stream, and learn specific categories' recent states.
%This process is illustrated by Table \ref{tbl:historyStates}.

We compare the topic coherence of history states with the topics learned from Wikipedia by LDA in terms of NPMI\cite{Rder2015ExploringTS}.
Different from traditional experiments that only compute the topic coherence of top words, we want to check whether  it can hold for more words.
Due to the limit of NPMI computing module\footnote{\url{https://github.com/AKSW/Palmetto}}, which computes the coherence of up to 10 words each time, we compute NPMI on the combination of top five words and each next five words as Table \ref{tbl:NPMIDetails}.

@@

\begin{table}[]
\centering
\scalebox{0.76}{
\begin{tabular}{|c|c|l|c|}
\hline
\#group* & \#word*  & words & NPMI\\ \hline
0 & 6-10 & airlines aviation flying pilot squadron &  0.113\\ \hline
1 & 11-15 &flights pilots raf airways fighter & 0.155\\ \hline
2 & 16-20 &boeing runway force crashed flew   & 0.092\\ \hline
3 & 21-25 &airfield landing passengers plane aerial & 0.179\\ \hline
4 & 26-30 &bomber radar wing bombers crash & 0.137\\ \hline
5 & 31-35 &airbus airports operations jet helicopter & 0.189\\ \hline
6 & 36-40 &squadrons base flown havilland crew & 0.088\\ \hline
7 & 41-45 &combat luftwaffe aerodrome carrier fokker & 0.159\\ \hline
8 & 46-50 &planes fly engine takeoff fleet & 0.186\\ \hline
9 & 51-55 &fuselage helicopters aviator naval aero & 0.157\\ \hline
10 & 56-60 &glider command training balloon faa & 0.166\\ \hline
\(\cdots\) & \(\cdots\) &\(\cdots\) &\(\cdots\)\\ \hline
18 & 96-100 &scheduled carriers military curtiss biplane &0.131\\ \hline
19 & 101-105 &accident engines iaf albatross rcaf &0.068\\ \hline
\end{tabular}
}
\caption{Top words' topic coherence of \textit{Aviation}'s history state. * means the group also contains the five top words \textit{aircraft}, \textit{air}, \textit{airport}, \textit{flight}, and \textit{airline}, but we don't show them in table to save space. NPMI is computed on ten words (a combination of words in each row and the five top words).}
\label{tbl:NPMIDetails}
\end{table}

\begin{figure}
        \centering
        \includegraphics[width=1.0\columnwidth]{img/NPMI.pdf}
        \caption{Effectiveness of history state in terms of NPMI}
        \label{fig:NPMI}
\end{figure}

@@

\begin{comment}
\begin{table*}[ht]
\centering
\caption{History States initialized from Wikipedia and topical words learned from HS-Prior-LDA}
\scalebox{0.76}{
\begin{tabular}{|cc|cc|cc|cc|cc|cc|}
\hline
\multicolumn{2}{|c|}{\textit{Aviation}} & \multicolumn{2}{c|}{\textit{Health}} & \multicolumn{2}{c|}{\textit{Middle East}} & \multicolumn{2}{c|}{\textit{Military}} & \multicolumn{2}{c|}{\textit{Mobile Phones}} & \multicolumn{2}{c|}{\textit{Video Games}} \\
\begin{tabular}[c]{@{}c@{}}Normal\\ States\end{tabular} & \begin{tabular}[c]{@{}c@{}}Topical\\ Words\end{tabular} & \begin{tabular}[c]{@{}c@{}}Normal\\ States\end{tabular} & \begin{tabular}[c]{@{}c@{}}Topical\\ Words\end{tabular} & \begin{tabular}[c]{@{}c@{}}Normal\\ States\end{tabular} & \begin{tabular}[c]{@{}c@{}}Topical\\ Words\end{tabular} & \begin{tabular}[c]{@{}c@{}}Normal\\ States\end{tabular} & \begin{tabular}[c]{@{}c@{}}Topical\\ Words\end{tabular} & \begin{tabular}[c]{@{}c@{}}Normal\\ States\end{tabular} & \begin{tabular}[c]{@{}c@{}}Topical\\ Words\end{tabular} & \begin{tabular}[c]{@{}c@{}}Normal\\ States\end{tabular} & \begin{tabular}[c]{@{}c@{}}Topical\\ Words\end{tabular} \\ 
\hline
aircraft & air\scriptsize(3253) & health & weight\scriptsize(16344) & al & \textcolor{red}{\textit{\#syria}}\scriptsize(4212) & army & killed\scriptsize(4055)  & android & iphone\scriptsize (13674) & game & games\scriptsize(8812) \\ 
air & plane & patients & loss & israel & \textcolor{red}{\textit{\#bahrain}} & military & news & mobile & apple & player & liked \\ 
airport & flight & medical & diet & iran & people & air & \textcolor{red}{\textit{\#libya}}\scriptsize(3503) & nokia & android & playstation & free \\ 
flight & time & disease & health & arab & israel & command & libya & ios & app & gameplay & xbox \\
airline & airlines & treatment & cancer & israeli & police & force & rebels & phone & ipad & nintendo & 360 \\
airlines & news & hospital & lose & egypt & \textcolor{red}{\textit{\#libya}}\scriptsize(2557) & regiment & people & samsung & samsung & games & playing \\
aviation & boat & patient & fat & egyptian & \#egypt & forces & police & game & mobile & players & played \\
flying & airport & clinical & tips & ibn & news & squadron & war & app & blackberry & xbox &iphone\scriptsize(2820)\\
pilot & force & symptoms & treatment & jerusalem & \#israel & infantry & libyan & iphone & tablet & mode & time\\
squadron & fly & cancer & body & syria & world & battle & attack & htc & apps & arcade & \textcolor{red}{\textit{ps3}}\\
\hline
\end{tabular}
}
\label{tbl:historyStates}
\end{table*}
\end{comment}
